---
title: "Fieldtrip2019_immune_data_worms"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Immune data from 2019 Field trip

Here are the data produced from the collaboration between the immunology FU and our group. I will be cleaning the data, reproducing the analysis done by fu and then followingly reproducing the results with the parasite Eimeria instead of worms.

#### Loading and cleaning the data
Libraries that I will need: 
```{r, echo=FALSE}
library(tidyverse)
library(tidyr)
library(readxl)
library(dplyr)
library(assertive)
library(stringr)
library(ggplot2)
```
Loading the data:
ID: Immune Data data frame
```{r, echo=FALSE}
ID <- read_excel("Mouse_Eimeria_Field/data_input/Fieldtrip2019datasheet_Hongwei20211122_edit_Fay.xlsx", sheet = 1)
```
Start cleaning:
```{r pressure, echo=FALSE}
# Start with viewing the structure of your data
glimpse(ID)
summary(ID)
```
Now go on with the cleaning
**1.Lines Spaces**
**2. Type of columns**
```{r}
assert_is_numeric()
bike_share_rides <- bike_share_rides %>%
  # Remove 'minutes' from duration: duration_trimmed
  mutate(duration_trimmed = str_remove(duration, "minutes"),
         # Convert duration_trimmed to numeric: duration_mins
         duration_mins = as.numeric(duration_trimmed))
```
**3. What about the ranges of each column? **
```{r}
# Create breaks
breaks <- c(min(bike_share_rides$duration_min), 0, 1440, max(bike_share_rides$duration_min))

# Create a histogram of duration_min
ggplot(bike_share_rides, aes(duration_min)) +
  geom_histogram(breaks = breaks)
```


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
