---
title: "Report: Hybrid vigor in response to Eimeria in the HMHZ"
author: "Alice"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  pdf_document:
    toc: yes
    toc_depth: '4'
    fig_caption: yes        
    includes:  
      in_header: preamble-latex.tex
    keep_tex: yes
  word_document:
    toc: yes
    toc_depth: 4
    fig_caption: yes        
    includes:  
      in_header: preamble-latex.tex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(ggmap)
library(data.table)
library(plyr)
library("VennDiagram")
library(grid)
library(gridExtra)
library(knitr)
source("functions/HMHZ_Functions.R")
source("functions/makeMiceTable.R")

## Load data from oocysts counting 
flotDF <- read.csv("../data/Eimeria_detection/FINALOocysts2015to2017.csv")
LorenzoDF <- read.csv("../data/Eimeria_detection/Eimeria_oocysts_2015&2017_Lorenzo.csv")
# ## Import PCR data
# PCRdf <- read.csv("../data/Eimeria_detection/Inventory_contents_all.csv")
# ## Import qPCR data
# qpcrData <- read.csv("../data/Eimeria_detection/FINALqpcrData_2016_2017_threshold5.csv")
# # qpcrData <- qpcrData[!is.na(qpcrData$delta_ct_ilwe_MminusE) & !is.na(qpcrData$delta_ct_cewe_MminusE),]

#################### Load full data ####################
# General data
miceTable <- read.csv("../data/MiceTable_fullEimeriaInfos_2014to2017.csv")

# Remove other rodents
miceTable <- miceTable[!miceTable$Species %in% "Pet mus musculus",]
miceTable <- miceTable[-grep("ZZ", miceTable$Mouse_ID),]
miceTable <- miceTable[!miceTable$Mouse_ID %in% paste0("A_000", 1:3),]
## Remove wildpark Schorfheide (not needed, test)
wsh <- c(paste0("AA_000", 1:9), paste0("AA_00", 10:46))
miceTable <- miceTable[!miceTable$Mouse_ID %in% wsh ,]

# Give a PCR status
# by default, I enter PCRstatus as negative, then overwrite
miceTable$PCRstatus = "negative"
  # PCR positive = one of the 3 other markers than AP5 sequenced 
# (Ap5 was used for detection only, the other markers for confirmation)
miceTable$PCRstatus[miceTable$n18S_Seq == "positive" | 
                      miceTable$COI_Seq == "positive" | 
                      miceTable$ORF470_Seq == "positive"] <- "positive"
 
# PCRstatus is NA if everything is NA
miceTable$PCRstatus[is.na(miceTable$n18S_Seq) &
                      is.na(miceTable$COI_Seq) &
                      is.na(miceTable$ORF470_Seq)] <- NA
```

# Eimeria detection oocysts flotation

## Improving Eimeria oocysts detection

```{r oocystsDetec, echo=FALSE, warning=FALSE}
flotDF$OPG <- as.numeric(as.character(flotDF$OPG))
flotDF <- flotDF[!is.na(flotDF$OPG),]
## Lorenzo count (in 1mL dilution) for comparison
LorenzoDF <- LorenzoDF[!is.na(LorenzoDF$OPG),]
### Plot comparative Alice (dilution 0.1mL for most samples) and Lorenzo (dilution 1mL)
compData <- merge(flotDF, LorenzoDF, by = "Mouse_ID", all = T)
### Comparison 2 methods of flotation
# How many samples new were detected by decreasing the dilution?
N1 <- sum(compData$OPG.x > 0 & compData$OPG.y == 0, na.rm = T)
adjrsq <- summary(lm(compData$OPG.x ~ compData$OPG.y))$adj.r.squared
ggplot(
  compData, aes(x = OPG.x+1, y = OPG.y+1)) +
  geom_point(alpha = .5, size = 4) +
  coord_equal(ratio=1) +
  xlab("OPG + 1 counted in 0.1ml") + 
  ylab("OPG + 1 counted in 1ml") + 
  geom_smooth(method = "lm", se = FALSE, col = "red") +
  geom_abline(intercept = 0, slope = 1, linetype = 3) +
  scale_y_log10() + 
  scale_x_log10() +
  theme_bw()
```

`r N1` new samples were detected while diluting by 0.1mL PBS instead of 1mL before counting in Neubauer chamber.

Adjusted R-squared = `r round(adjrsq, 2)` represents the amount of variation in y explained by x.

## OPG that we keep

Number of Mus musculus caught with OPG values: `r nrow(miceTable[!is.na(miceTable$OPG),])`

```{r oocystssmooth, echo=FALSE, warning=FALSE}
ggplot(miceTable[!is.na(miceTable$OPG) &
                               miceTable$OPG >0,], aes(x = HI, y = OPG+1)) +
  geom_point(aes(fill = as.factor(Year)), pch = 21, alpha = .8, size = 4) +
  geom_smooth(se=F) +
  scale_y_log10() +
  theme_bw() +
  theme(legend.position="top") +
  theme(legend.title = element_blank()) +
  facet_grid(. ~ Year)
```

# Eimeria detection PCR

PCR positive = one of the 3 other markers than AP5 sequenced (Ap5 was used for detection only, the other markers for confirmation)

```{r pcr, echo=FALSE, warning=FALSE}
# plot PCR
ggplot(miceTable[miceTable$PCRstatus == "positive",], 
       aes(y = HI, x = eimeriaSpecies)) +
  geom_violin() +
  geom_jitter(aes(fill = eimeriaSpecies),
              width = .1, pch = 21, alpha = .4, size = 4) +
  theme_bw() +
  theme(legend.position="top") +
  theme(legend.title = element_blank()) +
  ggtitle("Violin plots on PCR data")
```

PCR positive = one of the 3 markers 18S, COI or ORF470) gave a sequence. Number of Mus musculus caught with PCR performed: `r nrow(miceTable[!is.na(miceTable$n18S_Seq)|!is.na(miceTable$COI_Seq)|!is.na(miceTable$ORF470_Seq),])`

# Eimeria detection qPCR 

We keep only the values for mice having been tested for BOTH ileum and cecum!

```{r qpcr, echo= FALSE}
# plot qPCR
ferrisiDFqpcr <- miceTable[!is.na(miceTable$delta_ct_cewe_MminusE) &
                       miceTable$delta_ct_cewe_MminusE > -5 &
                         !is.na(miceTable$eimeriaSpecies) &
                         miceTable$eimeriaSpecies == "E_ferrisi",]

ggplot(ferrisiDFqpcr, 
       aes(x = HI, y = delta_ct_cewe_MminusE)) +
  geom_point() +
  facet_grid(.~eimeriaSpecies) +
  geom_smooth() +
  theme_classic() +
  ggtitle("ferrisiqpcr")

ggplot(miceTable[miceTable$Year %in% c(2016, 2017),],
       aes(x = HI, y = delta_ct_cewe_MminusE, col = OPG > 0)) +
  geom_point(size = 5) + 
  geom_hline(yintercept = -5) +
  facet_grid(.~Year) +
  theme_bw() +
  ggtitle("Remark of J. Wolinska: some individuals here HAVE qPCR value, but no oocyst count?")

# Todo later: test if points for HI ~ 0.5 are dead ends...
```

# General stats on sampling 

```{r generalstats, echo = FALSE, message=FALSE}
## Which mice are not found (no HI given for these mice)?
miceInfoNeeded <- miceTable$Mouse_ID[is.na(miceTable$HI)]
# latitude or longitude missing for mice:
latLongMissing <- miceTable$Mouse_ID[
  is.na(miceTable$Latitude) |
    is.na(miceTable$Longitude)]

# Total
Nmice <- nrow(miceTable)
Nfarm <- length(unique(miceTable$farm))

# Create map of samples
mapHMHZ <- HI.map(df = miceTable[!is.na(miceTable$Longitude) & !is.na(miceTable$Latitude), ], 
                  size = 2, alpha = .3, margin = 0.2, zoom = 8) 
mapHMHZ

# mean and 95% ci of N of mice caught / farm (assuming normal distribution)
MEAN <- mean(by(miceTable, miceTable["farm"], nrow))
CI <- qnorm(0.975)*sd(by(miceTable, miceTable["farm"], nrow))/
  sqrt(nrow(miceTable))

# Hybrid index calculation:
minHINloci = min(as.numeric(substr(miceTable$HI_NLoci, 4,6)), na.rm = T)
maxHINloci = max(as.numeric(substr(miceTable$HI_NLoci, 4,6)), na.rm = T)
meanHINloci = round(mean(as.numeric(substr(miceTable$HI_NLoci, 4,6)), na.rm = T))
```

* Some information regarding latitude and longitude are missing for the following mice:

`r latLongMissing`

* We still miss info (HI) on the following mice (ask Jarda):

`r miceInfoNeeded` 

# General informations on HMHZ

* `r Nmice` mice were captured over three years, from `r Nfarm` farms

* On average, `r round(MEAN,2)` mice were caught per farm (95% CI `r round(CI,2)`)

* **Hybrid indexes** were calculated as ratio of M.m.d/M.m.m alleles (between `r minHINloci` and `r maxHINloci`, on average `r meanHINloci` loci)

```{r plotDensHI, echo=FALSE, warning=FALSE, fig.width=5,fig.height=3,fig.cap="\\label{fig:plot1}Number of animals caught along the hybrid index "}
ggplot(miceTable, aes(x = HI)) +
  geom_histogram(binwidth = 0.05, col = "black", fill = "lightblue") +
  theme_bw() +
  ggtitle("Density of hybrids") +
  facet_grid(. ~ Year)
```

# Prevalence of our 3 different methods

## Prevalence tables

```{r prevcompared, echo = FALSE}
prevalenceFlotation <- getPrevalenceTable(table(miceTable$OPG > 0, 
                                                miceTable$Year))
prevalencePCR <- getPrevalenceTable(table(miceTable$PCRstatus == "positive", 
                                          miceTable$Year))
prevalenceqPCR <- getPrevalenceTable(table(miceTable$delta_ct_cewe_MminusE > - 5, 
                                           miceTable$Year))

# samples are negative appart if ONE at least is positive
miceTable$allDetectionMethod <- "negative"

miceTable$allDetectionMethod[miceTable$OPG > 0 |
                               miceTable$PCRstatus == "positive" |
                              miceTable$delta_ct_cewe_MminusE > - 5] <- "positive"

prevalenceTot <- getPrevalenceTable(table(miceTable$allDetectionMethod,
                                          miceTable$Year ))

```

```{r tabPrevOO, echo=FALSE, results='asis'}
library(knitr)
kable(prevalenceFlotation, caption = "Prevalence of Eimeria per year, based on oocyst flotation")
```

```{r tabPrevPCR, echo=FALSE, results='asis'}
library(knitr)
kable(prevalencePCR, caption = "Prevalence of Eimeria per year, based on PCR detection. A mouse was considered infected by Eimeria if one of the 3 markers (COI, 18S or ORF470) gave a sequence")
```

```{r tabPrevqPCR, echo=FALSE, results='asis'}
library(knitr)
kable(prevalenceqPCR, caption = "Prevalence of Eimeria per year, based on qPCR in cecum")
```

```{r tabPrevtot, echo=FALSE, results='asis'}
kable(prevalenceTot, caption = "Prevalence of Eimeria per year, based on all detections methods. A mouse was considered infected by Eimeria if one of the 3 markers (COI, 18S or ORF470) gave a sequence, OR if it had a positive count of oocysts in its feces, OR if it was qPCR positive in cecum tissue")
```

## OPG-PCR

```{r opgpcr, echo=FALSE, warning=FALSE}
# first, compare PCR and oocysts
completeData1 <- miceTable[!is.na(miceTable$OPG) &
                             !is.na(miceTable$PCRstatus),]

myVennDiagram2 <- function(data){      
  area1 = nrow(subset(data, PCRstatus == "positive"))
  area2 = nrow(subset(data, OPG > 0))
  ## areas of 2-group overlap
  cross.area = nrow(subset(data, PCRstatus == "positive" & OPG > 0))
  grid.newpage()
  draw.pairwise.venn(area1 = area1, 
                          area2 = area2, 
                          cross.area = cross.area, 
                          category = c("PCR", "OPG"),
                          col = "transparent", 
                          fill = c("grey","green"),
                          alpha = 0.50,                  
                          cex = 1.5, cat.cex = 1.5, fontfamily = "serif", fontface = "bold",
                          cat.col = c("grey","green"),
                          cat.fontfamily = "serif")
}

myVennDiagram2(completeData1)
```

## OPG-qPCR

```{r opgpcrVenn, echo=FALSE, warning=FALSE}
# first, compare PCR and oocysts
completeDataopgqpcr <- miceTable[!is.na(miceTable$OPG) &
                          !is.na(miceTable$delta_ct_cewe_MminusE),]

myVennDiagram3 <- function(data){      
  area1 = nrow(subset(data, delta_ct_cewe_MminusE > -5))
  area2 = nrow(subset(data, OPG > 0))
  ## areas of 2-group overlap
  cross.area = nrow(subset(data, delta_ct_cewe_MminusE > -5 & OPG > 0))
  grid.newpage()
  draw.pairwise.venn(area1 = area1, 
                     area2 = area2, 
                     cross.area = cross.area, 
                     category = c("qPCR", "OPG"),
                     col = "transparent", 
                     fill = c("grey","green"),
                     alpha = 0.50,                  
                     cex = 1.5, cat.cex = 1.5, fontfamily = "serif", fontface = "bold",
                     cat.col = c("grey","green"),
                     cat.fontfamily = "serif")
}

myVennDiagram3(completeDataopgqpcr)
```

```{r opgqpcr, echo=FALSE, warning=FALSE}
# data1 <- miceTable[!is.na(miceTable$delta_ct_cewe_MminusE) &
#                             !is.na(miceTable$OPG), ]
# ggplot(data1,
#        aes(x = delta_ct_cewe_MminusE, y = OPG)) +
#   geom_point(alpha = .5, size = 4, pch = 21) +
#   geom_smooth(method = "lm", se = FALSE, col = "red") +
#   theme_bw()  +
#   # scale_y_log10() +
#   ggtitle("Compare qPCR results and OPG")
# 
# summary(lm(data1$OPG ~ data1$delta_ct_MminusE))
```
## OPG-qPCR-PCR

```{r opgpcrpcr, echo=FALSE, warning=FALSE}
# to compare, keep only samples tested for the 3 methods
# completeData <- myData[!is.na(myData$OPG) &
#                                    !is.na(myData$PCRstatus) &
#                                    !is.na(myData$delta_ct_MminusE),]
# 
# myVennDiagram3 <- function(data){      
#   area1 = nrow(subset(data, PCRstatus == "positive"))
#   area2 = nrow(subset(data, OPG > 0))
#   area3 = nrow(subset(data, delta_ct_MminusE > - THRESHOLD))
#   ## areas of 2-group overlap
#   n12 = nrow(subset(data, PCRstatus == "positive" & OPG > 0)) 
#   n23 = nrow(subset(data, delta_ct_MminusE > - THRESHOLD & OPG > 0))
#   n13 = nrow(subset(data, PCRstatus == "positive" & delta_ct_MminusE > - THRESHOLD))
#   ## areas of 3-group overlap
#   n123 = nrow(subset(data, PCRstatus == "positive" & 
#                        OPG > 0 &
#                        delta_ct_MminusE > - THRESHOLD))
#   grid.newpage()
#   draw.triple.venn(area1 = area1, area2 = area2, area3 = area3,
#                    n12 = n12, n23 = n23, n13 = n13,
#                    n123 = n123,
#                    category = c("PCR", "OPG", "qPCR"),
#                    col = "transparent", 
#                    fill = c("grey","green","orange"),
#                    alpha = 0.50,                  
#                    cex = 1.5, cat.cex = 1.5, fontfamily = "serif", fontface = "bold",
#                    cat.col = c("grey","green","orange"),
#                    cat.fontfamily = "serif")
# }
# ```
# 
# ```{r venn2, echo=FALSE, warning=FALSE, message=FALSE, fig.width=5,fig.height=3,fig.cap="\\label{fig:venn1}Comparison of detection: PCR vs flotation vs qPCŔ"}
# myVennDiagram3(completeData)
```

# Testing hybrid vigor along HMHZ

```{r}
ggplot(miceTable, aes(x = eimeriaSpecies, y = OPG +1)) +
  geom_violin() +
  geom_jitter(width = .1, pch = 21, size = 2) +
  scale_y_log10() +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 

ferrisiDFflot <- miceTable[!is.na(miceTable$OPG) &
                         miceTable$OPG > 0 &
                         !is.na(miceTable$eimeriaSpecies) &
                         miceTable$eimeriaSpecies == "E_ferrisi",]

ggplot(ferrisiDFflot, aes(x = HI, y = OPG)) +
  geom_point() +
  scale_y_log10() +
  geom_smooth()

## qPCR by OPG
ferrisiDFflot_n_qpcr <- miceTable[!is.na(miceTable$OPG) &
                                    miceTable$OPG > 0 &
                                    !is.na(miceTable$delta_ct_cewe_MminusE) &
                                    miceTable$delta_ct_cewe_MminusE > -5 &
                                    !is.na(miceTable$eimeriaSpecies) &
                                    miceTable$eimeriaSpecies == "E_ferrisi",]
  
ggplot(ferrisiDFflot_n_qpcr, aes(x = delta_ct_cewe_MminusE, y = OPG)) +
  geom_point() +
  scale_y_log10() +
  geom_smooth(method = "lm") +
  theme_classic()

## BCI vs OPG
ggplot(ferrisiDFflot, aes(x = BCI, y = OPG)) +
  geom_point() +
  scale_y_log10() +
  geom_smooth(method = "lm") +
  theme_classic()

summary(lm(ferrisiDFflot$BCI ~ ferrisiDFflot$OPG))

## BCI vs qPCR
ggplot(ferrisiDFqpcr, 
       aes(x = BCI, y = delta_ct_cewe_MminusE)) +
  geom_point(aes(fill = HI), pch = 21, size = 5) +
  scale_fill_gradient(low = "blue", high = "red") +
  geom_smooth(method = "lm") +
  theme_classic()

summary(lm(ferrisiDFqpcr$BCI ~ ferrisiDFqpcr$delta_ct_cewe_MminusE))

hist(miceTable$BCI, breaks = 50) 

## BCI vs qPCR
ferrisiDFqpcr$delta_ct_cewe_MminusE

ferrisiDFqpcr$intensity <- "high"
ferrisiDFqpcr$intensity[ferrisiDFqpcr$delta_ct_cewe_MminusE < -2] <- "middle"
ferrisiDFqpcr$intensity[ferrisiDFqpcr$delta_ct_cewe_MminusE < -4] <- "low"

ggplot(ferrisiDFqpcr, 
       aes(x = HI, y = BCI)) +
  geom_point(aes(fill = delta_ct_cewe_MminusE), pch = 21, size = 5) +
  facet_grid(.~intensity) +
  geom_smooth() +
  theme_classic()

```


Discussed with Stuart:
  
  * Test distributions 0 or counts. Test all vs only infected ("intensity") distribution. We should be able to fit the distribution of infected on all. Zeros are data. Stochastic move.
* Separation of the zero class. balanced design case/control ~ 400 +/-70infectés SNPchip. 
* H0: no differences are observed
* Separate <0.5 and >0.5 to see the species effect
- timing : WHEN (for my thesis?)
