---
title: "Report: Hybrid vigor in response to Eimeria in the HMHZ"
author: "Alice"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  pdf_document:
    toc: yes
    toc_depth: '4'
    fig_caption: yes        
    includes:  
      in_header: preamble-latex.tex
  word_document:
    toc: yes
    toc_depth: 4
    fig_caption: yes        
    includes:  
      in_header: preamble-latex.tex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("../R/Data_Preparation_Alice.R")
```

# General informations on HMHZ

* `r sum(prevFarmYear$N)` mice were captured over three years and had fecal samples processed, from `r nrow(prevFarmYear)` farms. 

* From these mice, *(tbc : N had colon content and intestinal tissues collected for PCR and qPCR detection)* 

* `r round(mean(by(myData, myData["farm"], nrow)), 2)` mice were caught on average by farm (95% ci : `r round(qnorm(0.975)*sd(by(myData, myData["farm"], nrow))/sqrt(nrow(myData)), 2)`)

* **Hybrid indexes** were calculated as ratio of M.m.d/M.m.m alleles (between `r minHINloci` and `r maxHINloci`, on average `r meanHINloci` loci)

```{r plotDensHI, echo=FALSE, warning=FALSE, fig.width=5,fig.height=3,fig.cap="\\label{fig:plot1}Number of animals caught along the hybrid index "}
plotDensHI <- ggplot(myData, aes(x = HI)) +
  geom_histogram(binwidth = 0.05, col = "black", fill = "lightblue") +
  theme_bw()
plotDensHI
```

```{r map, echo=FALSE, warning=FALSE, fig.width = 6, fig.height = 4, fig.cap="\\label{fig:map}Samples map. Each point correspond to one location, a less pronounced transparency indicating more animals sampled at this location. Hybrid index is represented by a gradient from blue (M.m.d) to red (M.m.m)"}
mapHMHZ
```

The average *Eimeria* prevalence per farm based on oocysts flotation is `r round(mean(prevFarmYear$prev), 2 )`. We observed, based on this technique,  a variation between years (Table 1).

```{r tabPrev, echo=FALSE, results='asis'}
library(knitr)
kable(prevalenceFlotation, caption = "Prevalence of Eimeria based on OPG per year")
```

## Improving Eimeria oocysts detection

`r N1` new samples were detected while diluting by 0.1mL PBS instead of 1mL before counting in Neubauer chamber.

Adjusted R-squared = `r round(adjrsq, 2)` represents the amount of variation in y explained by x. 

*https://www.r-bloggers.com/correlation-and-linear-regression/ (for Lorenzo)*

```{r plot1, echo=FALSE, warning=FALSE,  fig.width=5,fig.height=3,fig.cap="\\label{fig:plot1}Comparison of OPG depending on dilution level. Red line represents linear relationship between both axis, dotted line represents the function y = x"}
plot1
```

## Missing data (to complete with Victor)

Some mice do not have an hybrid index yet: `r missingHIMice`

## PCR

```{r, echo=FALSE, results='asis'}
library(knitr)
kable(getPrevalenceTable(table(myData$Ap5_PCR, myData$year)), 
      caption = "Prevalence of Eimeria based on positive AP5 PCR, per year")
```

*PCR.positive* means that AP5 was positive AND another marker gave a sequence

```{r, echo=FALSE, results='asis'}
library(knitr)
kable(tabpcr,
      caption = "Prevalence of Eimeria based on the obtention of a sequence of a marker (COI, 18S or ORF470), per year")
```

## qPCR

```{r, echo=FALSE, results='asis'}
library(knitr)
kable(tabqpcr,
      caption = "Prevalence of Eimeria based on positive qPCR, per year")
```

## Comparison oocysts flotation, PCR, qPCR 

*Plot a venn diagram here*

# Testing hybrid vigor along HMHZ

## Oocyst shedding proxy

First approximation:

```{r plotSmoothOPG, echo=FALSE, warning=FALSE, fig.width=6, fig.height=4, fig.cap="\\label{fig:plotSmoothOPG}OPG along HI, colored per year. Blue line represent a smooth function (method = loess)"}
plotSmoothOPG
```

Statistical model (dvp...)

## qPCR proxy

tbc

## BCI proxy

First approximation:

```{r plotBCI, echo=FALSE, warning=FALSE, fig.width=6, fig.height=4, fig.cap="\\label{fig:plotSmoothOPG}BCI along HI, colored per level of OPG"}
plotBCI
```


